---
title: "Data manipulation"
output: html_document
---

```{r}
library(readr)
library(tidyverse)
```


Import FAS litters and pups.

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

```{r}
pups_df = 
  read_csv("data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
```

## `select` is used to select columns

```{r}
select(litters_df, group, litter_number, gd0_weight)

select(litters_df, group, gd0_weight:gd_of_birth)

select(litters_df, group, starts_with("pups"))

select(litters_df, -litter_number)

select(litters_df, -starts_with("gd"))

select(litters_df, group, litter_id = litter_number)

select(litters_df, group, litter_id = litter_number, everything())
```

# `select` vs `pull`
# select will always give you a data frame; output is a dataframe
# pull - extract column from data frame, transforms it into a vector, so it's no longer in your dataframe; output is a vector; use pull instead of a $

```{r}
select(litters_df, group)
pull(litters_df, group)
```

# `rename` function
rename(litters_df, litter_id = litter_number)

# `relocate` function
```{r}
relocate(litters_df, litter_number)
```

# Exercise with pups_df
```{r}
colnames(pups_df)
pups_subset <- select(pups_df, litter_number, sex, pd_ears)
```

## `filter`

get rid of rows using `filter`

```{r}
filter(litters_df, group == "Mod8")

filter(litters_df, group != "Con7")

filter(litters_df, gd0_weight > 20)

filter(litters_df, group == "Con7" | group == "Con8")

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, !(group == "Con7"))

filter(litters_df, group == "Con7" | group == "Con8", gd0_weight > 20)

```

## `mutate`

this is used to add or change variables.

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)

mutate(litters_df, group = str_to_lower(group))

mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

## Missing values

```{r}
drop_na(litters_df)
drop_na(litters_df, gd0_weight)

```

## `arrange`

Can arrange by one or multiple columns.

```{r}
arrange(litters_df, gd0_weight)
arrange(litters_df, group, gd0_weight)
```


## pipes!
```{r}
litter_df = 
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |> 
  select(-starts_with("pups")) |> 
  drop_na(gd0_weight)
```




