---
title: "Simple document"
output: github_document
---


Let's import the 'FAS_litters.csv' csv using a relative path. Tidyverse using readr package (use for csv files). Readxl use for excel files.


```{r}
library(tidyverse)
library(readxl)
library(haven)
```

```{r}
litters_df =
  read_csv("data/FAS_litters.csv")

litters_df =
  janitor::clean_names(litters_df)
```

Let's import the 'FAS_pups.csv' csv using a relative path. Janitor clean names makes all variable names snakecase.

```{r}
pups_df = 
  read_csv("data/FAS_pups.csv")

pups_df = 
  janitor::clean_names(pups_df)

names(pups_df)
```


## look at data

```{r}
litters_df

head(litters_df)

tail(litters_df)
```

YOu can use view. Pull up separate window too look at dataset.

```{r, eval = FALSE}
view(litters_df)
```


Look at a data summary. Structure tells you column, size, col types. Skim has neat summary of data; helpful to skim looking at dataset for first time.
```{r}
str(litters_df)

skimr::skim(litters_df)
```

Practice with read_*
```{r}
litters_df = 
  read_csv("data/FAS_litters.csv",
           skip = 10, col_names = FALSE)
```

Look at NA values
```{r}
litters_df = 
  read_csv(
    "data/FAS_litters.csv",
    na = c("NA", 19)
  )
```

Column types
```{r}
litters_df = 
  read_csv("data/FAS_litters.csv",
           col_types = 
             cols(
               Group = col_factor(),
               `GD0 weight` = col_character()
             ))
```

##Import other types of files.
Import a xlsx file first.

```{r}
mlb_df = 
  read_excel("data/mlb11.xlsx")
```

Import a SAS file.
```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat")
```

##Exporting data
We have code that cleans data and need to export it.

```{r}
litters_df_cleaned = 
  read_csv("data/FAS_litters.csv")

litters_df_cleaned = 
  janitor::clean_names(litters_df_cleaned)

write_csv(litters_df_cleaned,"data/litters_cleaned.csv")
```


